---
title: "Paper Result Analysis"
author: "Benjamin Grier"
format: 
  html:
    toc: true
    code-links:
      - text: Analysis and plotting functions
        icon: file-code
        href: https://github.com/BG-AIMS/kingfish-climate-risk/blob/main/mCRS%20pipeline%20Scripts/results_analysis_functions.R
    embed-resources: true
    
editor: visual
---

Climate Risk Assessment for Yellowtail Kingfish - Statistical Analysis

## Load Required files and setup analysis file

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'c:/Users/grier/Documents/Projects/')
#setwd('/home/ben/Documents/Projects/') # Set working directory to include all required files.
library(data.table)
library(ggplot2)
library(sf)
library(ggpubr)
library(maps)
library(dplyr)
options(dplyr.summarise.inform = FALSE)
library(scico)
library(nlme)
library(emmeans)
library(SIBER)
library(distfree.cr)
```

Load data files

```{r loading data files}
coordinates = read.csv("kingfish_data_dir/Required Files/Global_Coordinates.csv") # Load coordinate point data 
# Load results file for population level analyses for all depths
pop_all_depth_results <- read.csv('kingfish_data_dir/Results/mCRS-DeltamCRS RAW/DeltaCRS_mCRS_RAW_depth_together.csv')
# Load results file for species level analyses for all depths
species_results = read.csv('kingfish_data_dir/Results/mCRS-DeltamCRS RAW/DeltaCRS_mCRS_RAW_depth_together_SpL.csv')
species_results$DeltaCRS = as.numeric(species_results$DeltaCRS)
# Load results file for population level analyses for surface depth
surface_results = read.csv("kingfish_data_dir/Results/mCRS-DeltamCRS RAW/DeltaCRS_mCRS_RAW_surface_depth.csv")

# Load environmental data 
load('kingfish_data_dir/Required Files/historical_stack_ensemble.RData')
load('kingfish_data_dir/Required Files/ssp26_stack_ensemble.RData')
load('kingfish_data_dir/Required Files/ssp45_stack_ensemble.RData')
load('kingfish_data_dir/Required Files/ssp85_stack_ensemble.RData')
ENV <- list(historical.stack.ensemble, ssp26.stack.ensemble, ssp45.stack.ensemble, ssp85.stack.ensemble)
ENV.summary <- rbindlist(ENV) #combine all data tables into one ENV
rm(ENV,historical.stack.ensemble, ssp26.stack.ensemble, ssp45.stack.ensemble,ssp85.stack.ensemble, DataCollation)
ENV.summary <- ENV.summary[!ENV.summary$Extract.ID %in% c('103', '204', '200', '111', '112','34', '52'),]
ENV.summary.Y <- ENV.summary[,.(no3.mean = mean(no3), ph.mean = mean(ph), so.mean= mean(so), thetao.mean = mean(thetao)),by =.(Extract.ID,PopID,Year,Scenario,Depth)]
ENV.PopSummary <- ENV.summary.Y[,.(no3.mean = mean(no3.mean), ph.mean = mean(ph.mean), so.mean= mean(so.mean), thetao.mean = mean(thetao.mean)),by =.(PopID,Year,Scenario,Depth)]
ENV.PopSummary <- data.frame(ENV.PopSummary)
ENV.PopSummary_long <- tidyr::pivot_longer(ENV.PopSummary, cols = c(5:8), names_to = c('Variable'), values_to = 'Values')

# Load INLA spatio-temporal model results
load("kingfish_data_dir/Required Files/INLA_Res_Combined.RData")
```

Initial data processing of results files

```{r initial processing}
source("kingfish-climate-risk/mCRS pipeline Scripts/results_analysis_functions.R")
pop_all_depth_results = preprocess_results_df(pop_all_depth_results, coordinates, "population_all_depths")

species_results = preprocess_results_df(species_results, coordinates, "species")

surface_results = preprocess_results_df(surface_results, coordinates, "surface_depth")

combined_results = rbind(pop_all_depth_results, species_results, surface_results)
```

## Exploratory analysis of end-century results

::: panel-tabset
## CRS

```{r exploratory CRS}
ggplot() + geom_boxplot(data = combined_results[combined_results$Year == 2100,], aes(x = Depth, y = CRS, color = Level)) + facet_wrap(~PopLab)
```

## ðš«CRS

```{r exploratory DeltaCRS}
ggplot() + geom_boxplot(data = combined_results[combined_results$Year == 2100,], aes(x = Depth, y = DeltaCRS, color = Level)) + facet_wrap(~PopLab)
```
:::

## Identifying population-depth warming levels

To ensure result data are compared across consistent levels of future change results are calculated relative to local population-depth-level warming levels rather than a year value. Results values are reported and analysed at the year when a given population's temperature conditions reach a particular warming level for a future climate scenario. Therefore, results for each population are compared when that population and depth reach the same level of physical climate change impact, but this may not be at the same time.

### Calculate Warming Levels

```{r warming levels}
# Mean temperature levels 
meanTemp <- ENV.PopSummary[,c("PopID","Year","Scenario","Depth","thetao.mean")]
meanTemp = mean_Delta_temperature(meanTemp)

# Applying temperature GAM models 
newTempGAM <- meanTemp[,c("PopID","Depth","Year","Scenario")]
newTempGAM$DeltaTModelled <- 0
ssps = list('ssp26', 'ssp45', 'ssp85')
depths = list(10,50,100,150)
pops = list('NE_Pacific','NW_Pacific','South_Africa','South_Pacific')

## Population Specific Warming levels --
for(i in 1:length(ssps)){
  for(j in 1:length(depths)){
    for(z in 1:length(pops)){
      model <- mgcv::gam(DeltaTemp ~ s(Year), data = meanTemp[meanTemp$PopID == pops[z]&
                                                                meanTemp$Depth == depths[j] &
                                                                meanTemp$Scenario == ssps[i],])
      newTempGAM[newTempGAM$Scenario == ssps[i]&
                   newTempGAM$PopID == pops[z]&
                   newTempGAM$Depth == depths[j],]$DeltaTModelled <- predict(model,
                                                                             newTempGAM[newTempGAM$Scenario == ssps[i]&
                                                                                       newTempGAM$PopID == pops[z]&
                                                                                       newTempGAM$Depth == depths[j],])
    }
  }
}

# Find years when warming levels are reached for a given PopID and Depth 
ssp26.0.75deg.Year <- as.data.frame(newTempGAM[newTempGAM$Scenario == 'ssp26',]) %>% filter(DeltaTModelled >= 0.75) %>% 
  group_by(PopID, Depth, Scenario) %>% 
  summarize(Warming.year = min(Year))
ssp26.0.75deg.Year$Warming.level <- 0.75
ssp45.1deg.Year <- as.data.frame(newTempGAM[newTempGAM$Scenario == 'ssp45',]) %>% filter(DeltaTModelled >= 1) %>% 
  group_by(PopID, Depth, Scenario) %>% 
  summarize(Warming.year = min(Year))
ssp45.1deg.Year$Warming.level <- 1
ssp85.1.75deg.Year <- as.data.frame(newTempGAM[newTempGAM$Scenario == 'ssp85',]) %>% filter(DeltaTModelled >= 1.75) %>% 
  group_by(PopID, Depth, Scenario) %>% 
  summarize(Warming.year = min(Year))
ssp85.1.75deg.Year$Warming.level = 1.75
Warming.levels <- rbind(ssp26.0.75deg.Year, ssp45.1deg.Year, ssp85.1.75deg.Year)
Warming.levels$Depth <- as.factor(Warming.levels$Depth)

# Filter results to include warming level years only
warmlevel_data_all_depths <- left_join(pop_all_depth_results[!pop_all_depth_results$Scenario == 'historical',], Warming.levels, by = c('PopID','Depth','Scenario'))
warmlevel_data_all_depths <- warmlevel_data_all_depths[warmlevel_data_all_depths$Year == warmlevel_data_all_depths$Warming.year,]

warmlevel_data_surface <- left_join(surface_results[!surface_results$Scenario == 'historical',], Warming.levels, by = c('PopID','Depth','Scenario'))
warmlevel_data_surface <- warmlevel_data_surface[warmlevel_data_surface$Year == warmlevel_data_surface$Warming.year,]
```

### Warming levels figures

```{r warming level figures}
##### modelled temperature plots ----
meanTemp$Depth <- as.factor(meanTemp$Depth)
# axes ranges
minTemp <- min(newTempGAM$DeltaTModelled)
maxTemp <- max(newTempGAM$DeltaTModelled)
#Pop names to labels for plotting
Warming.levels$PopLab = Warming.levels$PopID
Warming.levels = PopID_add_PopLab(Warming.levels)

newTempGAM$PopLab = newTempGAM$PopID
newTempGAM = PopID_add_PopLab(newTempGAM)

meanTemp$PopLab = meanTemp$PopID
meanTemp = PopID_add_PopLab(meanTemp)
```

::: panel-tabset
## SSP1-2.6

```{r Warming Levels SSP1-2.6}
ggplot() +
  geom_line(data = newTempGAM[newTempGAM$Scenario == 'ssp26',], aes(x = Year, y = DeltaTModelled, color = as.factor(Depth))) +
  geom_line(data = meanTemp[meanTemp$Scenario == 'ssp26',],aes(x = Year, y = DeltaTemp, color = as.factor(Depth)), alpha = 0.2) +
  geom_segment(data = Warming.levels[Warming.levels$Scenario == 'ssp26',],aes(x = Warming.year,y = minTemp,xend = Warming.year, yend = Warming.level), alpha = 0.4) + 
  geom_segment(data = Warming.levels[Warming.levels$Scenario == 'ssp26'&Warming.levels$Depth ==150,],aes(x = 2001, y = Warming.level, xend = Warming.year, yend = Warming.level), alpha = 0.4) +
  geom_point(data = Warming.levels[Warming.levels$Scenario == 'ssp26',], aes(x = Warming.year, y = Warming.level), alpha = 0.5, size = 2) +
  scale_y_continuous(limits = c(minTemp, maxTemp)) + labs(color = 'Depth') + ylab('\u0394Temp (\u00b0C)') +
  facet_wrap(~PopLab, nrow = 1) + theme_classic() + ggtitle('Population Temperature change - SSP1-2.6') + 
  theme(axis.text = element_text(size = 9), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```

## SSP2-4.5

```{r Warming Levels SSP2-4.5}
ggplot() +
  geom_line(data = newTempGAM[newTempGAM$Scenario == 'ssp45',], aes(x = Year, y = DeltaTModelled, color = as.factor(Depth))) +
  geom_line(data = meanTemp[meanTemp$Scenario == 'ssp45',],aes(x = Year, y = DeltaTemp, color = as.factor(Depth)), alpha = 0.2) +
  geom_segment(data = Warming.levels[Warming.levels$Scenario == 'ssp45',],aes(x = Warming.year,y = minTemp,xend = Warming.year, yend = Warming.level), alpha = 0.4) + 
  geom_segment(data = Warming.levels[Warming.levels$Scenario == 'ssp45'&Warming.levels$Depth ==150,],aes(x = 2001, y = Warming.level, xend = Warming.year, yend = Warming.level), alpha = 0.4) +
  geom_point(data = Warming.levels[Warming.levels$Scenario == 'ssp45',], aes(x = Warming.year, y = Warming.level), alpha = 0.5, size = 2) +
  scale_y_continuous(limits = c(minTemp, maxTemp)) + labs(color = 'Depth') + ylab('\u0394Temp (\u00b0C)') +
  facet_wrap(~PopLab, nrow = 1) + theme_classic() + ggtitle('Population Temperature change - SSP2-4.5') + 
  theme(axis.text = element_text(size = 9), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```

## SSP5-8.5

```{r Warming Levels SSP5-8.5}
ggplot() +
  geom_line(data = newTempGAM[newTempGAM$Scenario == 'ssp85',], aes(x = Year, y = DeltaTModelled, color = as.factor(Depth))) +
  geom_line(data = meanTemp[meanTemp$Scenario == 'ssp85',],aes(x = Year, y = DeltaTemp, color = as.factor(Depth)), alpha = 0.2) +
  geom_segment(data = Warming.levels[Warming.levels$Scenario == 'ssp85',],aes(x = Warming.year,y = minTemp,xend = Warming.year, yend = Warming.level), alpha = 0.4) + 
  geom_segment(data = Warming.levels[Warming.levels$Scenario == 'ssp85'&Warming.levels$Depth ==150,],aes(x = 2001, y = Warming.level, xend = Warming.year, yend = Warming.level), alpha = 0.4) +
  geom_point(data = Warming.levels[Warming.levels$Scenario == 'ssp85',], aes(x = Warming.year, y = Warming.level), alpha = 0.5, size = 2) +
  scale_y_continuous(limits = c(minTemp, maxTemp)) + labs(color = 'Depth') + ylab('\u0394Temp (\u00b0C)') +
  facet_wrap(~PopLab, nrow = 1) + theme_classic() + ggtitle('Population Temperature change - SSP5-8.5') + 
  theme(axis.text = element_text(size = 9), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```
:::

## Comparing genetic-population and species level analyses

Climate risk assessments are often performed at a species-level, without considering genetic differentiation between populations of the same species. Genetically and geographically distinct populations are likely to experience different levels of climate risk because they are adapted to distinct climate niches (Li L. et al., 2022; Smith et al., 2019; Angert et al., 2011).

```{r genetic-population vs species level}
min_comb_crs = min(combined_results[
  combined_results$Level %in% c("population_all_depths", "species") &
  combined_results$Year == 2100,
  "CRS"
  ]
)
max_comb_crs = max(combined_results[
  combined_results$Level %in% c("population_all_depths", "species") &
  combined_results$Year == 2100,
  "CRS"
  ]
)
```

::: panel-tabset
## SSP1-2.6

```{r species comparison SSP1-2.6}
ggplot() + geom_boxplot(data = combined_results[combined_results$Scenario == 'ssp26'&combined_results$Year == 2100 & combined_results$Level %in% c("population_all_depths", "species"),], aes(x = Depth, y = CRS, fill = Level)) + 
  scale_y_continuous(limits = c(min_comb_crs, max_comb_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Niche Analysis Level - SSP1-2.6') + ylab('mCRS') +
  labs(fill = 'Analysis Level') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```

## SSP2-4.5

```{r species comparison SSP2-4.5}
ggplot() + geom_boxplot(data = combined_results[combined_results$Scenario == 'ssp45'&combined_results$Year == 2100 & combined_results$Level %in% c("population_all_depths", "species"),], aes(x = Depth, y = CRS, fill = Level)) + 
  scale_y_continuous(limits = c(min_comb_crs, max_comb_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Niche Analysis Level - SSP2-4.5') + ylab('mCRS') +
  labs(fill = 'Analysis Level') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```

## SSP5-8.5

```{r species comparison SSP5-8.5}
ggplot() + geom_boxplot(data = combined_results[combined_results$Scenario == 'ssp85'&combined_results$Year == 2100 & combined_results$Level %in% c("population_all_depths", "species"),], aes(x = Depth, y = CRS, fill = Level)) + 
  scale_y_continuous(limits = c(min_comb_crs, max_comb_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Niche Analysis Level - SSP5-8.5') + ylab('mCRS') +
  labs(fill = 'Analysis Level') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```
:::

:::{.panel-tabset}
## SSP1-2.6

```{r species comparison - aggregated SSP1-2.6}
ggplot() + geom_boxplot(data = combined_results[combined_results$Scenario == 'ssp26'&combined_results$Year == 2100 & combined_results$Level %in% c("population_all_depths", "species"),], aes(x = Level, y = CRS, fill = Level)) + 
  scale_y_continuous(limits = c(min_comb_crs, max_comb_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Niche Analysis Level - SSP1-2.6') + ylab('mCRS') +
  labs(fill = 'Analysis Level') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```

## SSP2-4.5

```{r species comparison - aggregated SSP2-4.5}
ggplot() + geom_boxplot(data = combined_results[combined_results$Scenario == 'ssp45'&combined_results$Year == 2100 & combined_results$Level %in% c("population_all_depths", "species"),], aes(x = Level, y = CRS, fill = Level)) + 
  scale_y_continuous(limits = c(min_comb_crs, max_comb_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Niche Analysis Level - SSP2-4.5') + ylab('mCRS') +
  labs(fill = 'Analysis Level') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```

## SSP5-8.5

```{r species comparison - aggregated SSP5-8.5}
ggplot() + geom_boxplot(data = combined_results[combined_results$Scenario == 'ssp85'&combined_results$Year == 2100 & combined_results$Level %in% c("population_all_depths", "species"),], aes(x = Level, y = CRS, fill = Level)) + 
  scale_y_continuous(limits = c(min_comb_crs, max_comb_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Niche Analysis Level - SSP5-8.5') + ylab('mCRS') +
  labs(fill = 'Analysis Level') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) 
```
:::

These plots show that climate risk scores are higher and have greater variation when populations are assessed separately as genetically-distinct populations rather than as a species. This is due to the reduced environmental conditions that the historical niche of each population incorporates compared to the historical niche of the species as a whole. If *S. lalandi* were assessed at a species level only, future climate risk levels may be underestimated, particularly under more severe future climate scenarios.

## Comparing adult-all depths and dispersal-surface life stages (at warming level years)

```{r comparing all depths and surface results}
depth_comparison_data = rbind(warmlevel_data_all_depths, warmlevel_data_surface)

min_depth_crs = min(depth_comparison_data$CRS)
max_depth_crs = max(depth_comparison_data$CRS)
```

::: panel-tabset
## SSP1-2.6

```{r surface comparison SSP1-2.6}
ggplot() + geom_boxplot(data = depth_comparison_data[depth_comparison_data$Scenario == 'ssp26',], aes(x = Depth, y = CRS, fill = Level)) + 
  #scale_y_continuous(limits = c(min_depth_crs, max_depth_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Life Stage Depths - SSP1-2.6') + ylab('mCRS') +
  labs(fill = 'Life Stage Depths') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) + facet_wrap(~PopID) 
```

## SSP2-4.5

```{r surface comparison 2-4.5}
ggplot() + geom_boxplot(data = depth_comparison_data[depth_comparison_data$Scenario == 'ssp45',], aes(x = Depth, y = CRS, fill = Level)) + 
  #scale_y_continuous(limits = c(min_depth_crs, max_depth_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Life Stage Depths - SSP2-4.5') + ylab('mCRS') +
  labs(fill = 'Life Stage Depths') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) + facet_wrap(~PopID) 
```

## SSP5-8.5

```{r surface comparison 5-8.5}
ggplot() + geom_boxplot(data = depth_comparison_data[depth_comparison_data$Scenario == 'ssp85',], aes(x = Depth, y = CRS, fill = Level)) + 
  #scale_y_continuous(limits = c(min_depth_crs, max_depth_crs), breaks = c(-4,-3,-2,-1,0,1)) +
  theme_classic() + ggtitle('Life Stage Depths - SSP5-8.5') + ylab('mCRS') +
  labs(fill = 'Life Stage Depths') + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 14, face = 'bold'), 
        strip.text.x = element_text(size = 10)) + facet_wrap(~PopID) 
```
:::

The substantially higher climate risk scores when surface depths are analysed separately, compared with all populations being assessed suggests that the *Seriola lalandi* egg-dispersal life stage is most at risk. *S. lalandi* eggs are positively bouyant (Moran et al., 2007), putting this life stage at greater risk.

The large differences in analysis results is due to the combination of depth levels in the `population_all_depths` analysis. Combining the depths together leads to the strong future nitrate changes at deep depths (100-150m) masking the surface changes. This leads to the populations experiencing the greatest climate risk/magnitudes of climate change at deep depths. However, when surface data is assessed separately (as a proxy for the conditions *S. lalandi* eggs may experience) the high level of climate impact on dispersal life stages becomes clear.

## Geographic patterns of Climate Risk Scores

```{r Geographic patterns of CRS}
all_depths_sf <- st_as_sf(warmlevel_data_all_depths, coords = c(8,9), crs = 4326) # make sf object to use in DeltaCRS mapping 

minDcrs = min(all_depths_sf$DeltaCRS)
maxDcrs = max(all_depths_sf$DeltaCRS)

BB_list = list(list(rbind(c(50,-15), c(50,-40), c(0,-40), c(0,-15), c(50,-15))),
               list(rbind(c(285,10), c(200,10), c(200, 40), c(285,40), c(285,10))),
               list(rbind(c(115,25), c(150,25), c(150, 45), c(115,45), c(115,25))),
               list(rbind(c(100,-5),c(100,-60),c(300,-60),c(300,-5),c(100,-5))))
Boxes <- st_multipolygon(BB_list)
Boxes <- st_sfc(Boxes, crs = 4326)
Boxes_360 <- st_shift_longitude(Boxes)
Warmlevel.data_360 <- st_shift_longitude(all_depths_sf)
World <- st_read('kingfish_data_dir/Required Files/GIS/ne_10m_land.shp', quiet=TRUE)
World_360 <- sf::st_as_sf(map('world2', plot = FALSE, fill = TRUE))

Test <- ggplot() + 
  geom_sf(data = World_360) +
  geom_sf(data =  Warmlevel.data_360[Warmlevel.data_360$Depth == 10 & Warmlevel.data_360$Scenario == 'ssp85',], aes(color = DeltaCRS), size = 2, alpha = 0.6) +
  theme_void() + theme(legend.position = "right") + 
  scale_color_scico(palette = 'batlow', midpoint = 0, limits = c(minDcrs, maxDcrs)) + labs(color = '\u0394mCRS') + 
  theme(legend.position = 'bottom')
leg <- ggpubr::get_legend(Test)
```

::: panel-tabset
## SSP1-2.6

```{r DeltaCRS Map SSP1-2.6}
plot(DeltamCRSMap(Scenario = "ssp26"))
```

## SSP2-4.5

```{r DeltaCRS Map SSP2-4.5}
plot(DeltamCRSMap(Scenario = "ssp45"))
```

## SSP5-8.5

```{r DeltaCRS Map SSP5-8.5}
plot(DeltamCRSMap(Scenario = "ssp85"))
```
:::

Differences in ðš«CRS between locations increase as the climate scenario increases in severity.

Main regions of low ðš«CRS compared to surrounding regions are:

-   Northern extents of the Australian range
-   North-East Pacific population
-   Eastern extents of the South African population

Main regions of high ðš«CRS compared to surrounding regions are:

-   Northern New Zealand ranges
-   South-Eastern Australian ranges
-   South Pacific locations

## Exploratory PCA plots

```{R PCA plots}
env_startcol=9
env_endcol=12
refyear_start="1850-01-01"
refyear_end="2000-01-01"
scenario="ssp85"
```

::: panel-tabset
## All Depths Together

```{r PCA All Depths Together}
plot(PCA_plotting(
  coordinates[,c(9,4)],
  ENV.summary[ENV.summary$Scenario %in% c("historical", scenario),],
  scenario, env_startcol, env_endcol, refyear_start, refyear_end,
  "All depths together PCA"
))
```

## Surface Level

```{r PCA Surface Level}
plot(PCA_plotting(
  coordinates[,c(9,4)],
  ENV.summary[ENV.summary$Scenario %in% c("historical", scenario) & ENV.summary$Depth == 10,],
  scenario, env_startcol, env_endcol, refyear_start, refyear_end,
  "Surface level PCA",
  method="surface"
))
```
:::

The PCA values (data from which CRS are calculated) when assessing all depths together show that the main change in conditions is towards greater nitrate concentration and salinity (i.e. conditions are moving towards the bottom right of the plot).

However, when assessing the surface level separately the PCA values show the main change in conditions occurring towards greater temperature and lower pH (sea surface warming and acidification). These are key environmental shifts that are likely to affect egg-dispersal life stages disproportionately compared to other life stages such as juvenile or adult *S. lalandi*, that move across a depth range.
